---
title: "fpp2_4"
author: "yinshu zhang"
date: "September 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
```

- names for `y`: regressand, dependent, explained variable, or forecast in this book. 
- names for `x`: regressors, independent, explanatory variables, in this book call predictor.

# simple linear regression

example:

```{r lm}
uschange %>% as.data.frame() %>% GGally::ggpairs() 
dat <- uschange[, c("Income","Consumption")]
autoplot(dat)
lm(Consumption ~ Income, data=as.data.frame(dat))
tslm(Consumption ~ Income, data=dat)
```

*without season and trend, `tslm` is the same as `lm`*

### a multi-lineare regression model

if $\beta$ are coefficients e is error, then $y_t=\beta_0+\beta_1x_{1,t}+\beta_2x_{2,t}+\beta_3x_{3,t}+e_t$

```{r fit}
fit.consMR <- tslm(
  Consumption ~ Income + Production + Unemployment + Savings,
  data=uschange)
summary(fit.consMR)
```

### fitted value
predciton of yt within training set.

$\hat{y}_t = \hat{\beta_0} + \hat{\beta_1}x_{1,t} + \hat{\beta_2}x_{2,t} + ... \hat{\beta_k}x_{k,t} $

### goodness-of-fit, R-squared
observed y and predicted $\hat{y}$

$R^2=\sum(\hat{y}_t - \bar{y})^2 / \sum(y_t - \bar{y})^2$

$R^2$ should close to 1 if prediciton is close to actual value, $R^2$ is between 0 and 1. Since this is base on training set, by adding more data, $R^2$ will never reduced, a better way is to validate with test set.

maximising $R^2$ is equivalent to minimising SSE.

### Adjuected R-squared
$\bar{R}^2 = 1-(1-R^2)(T-1)/(T-k-1)$ T is number of observations and k is the number of perdictors. best model has largest value of "R-var-squared".

### standard error of regression

standard deviation of residuals, or "redisual standard error".

# TS dummy variable

remember dummy variables need n-1 add to lm model, because one is intercept already.

example of quarters as dummy variable

```{r dummy}
beer2 <- window(ausbeer, start=1992)
fit.beer <- tslm(beer2 ~ trend + season)
summary(fit.beer)
```

## Intervention variables
adhoc unpredictable events

- if effect for one period, we use "spike variable", this is equivalent to dummary variable handlning an outlier.
- interventions have an immediate and permanent effect. If an intervention causes a level shift immediatly and permanently, we use "step" vaiable, which is, zero before and 1 afterwards
- Another form of permanent effect is a change of slope. Here the intervention is handled using a piecewise linear trend;

# Fourier series

An alternative to using seasonal dummy variables, especially for long seasonal periods, is to use Fourier terms. the advenagte is fewer variables.

> a series of sine and cosine terms of the right frequencies can approximate any periodic function

K is season/2

```{r forier}
fourier.beer <- tslm(beer2 ~ trend + fourier(beer2, K=2))
summary(fourier.beer)
```



